---
title: "p8105_hw2_yy2912.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
```
## problem 1
## Load in the trash data

```{r}
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel",
                 range= cell_cols("A2:N408"))%>%
                janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls, digit=0), sports_balls_new=as.integer(sports_balls))
```
## read and clean precipitation data from 2017 and 2018

```{r}
prec_data_2017=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate("year"='2017')%>%
  drop_na()

```

```{r}
prec_data_2018=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate("year"='2018')%>%
  drop_na()

```


```{r}
  prec_data_new = 
  bind_rows(prec_data_2017, prec_data_2018) %>%
  mutate(month = month.name [month] )
```



The number of observations in the dataset (trash_data) is `r nrow (trash_data)`

The number of observations in the dataset (prec_data_new) is `r nrow (prec_data_new)` 

the key variables in the dataset 

the total precipitation in 2018 is `r sum(pull(prec_data_2018,total))`

the median number of sports balls in a dumpster in 2017 is `r median(pull(trash_data,sports_balls))`

## problem 2
## Load in the fivethrityeight data

```{r}
polsmonth_data = read_csv(file = "./data/pols-month.csv")%>% 
separate(mon, into=c("year", "month", "day"))%>% 
mutate(month_numeric=as.numeric(month),month_pols=month.name[month_numeric],president='gop','dem')%>% select(-prez_dem, -prez_gop, -day)

```

```{r}
snp_data = read_csv(file = "./data/snp.csv")%>% 
separate(date, into=c("year", "month", "day"))%>%
mutate(month=month.name[month_numeric=as.numeric(month)])%>% 
select(-day)%>%
select("year", "month","close")
```

```{r}
unemployment_data = read_csv(file = "./data/unemployment.csv")%>% 
    janitor::clean_names() %>% 
  mutate(year=as.character(year))

```  
```{r}
unemployment_new_data = 
pivot_longer(unemployment_data, jan:dec,names_to = "month", values_to = "%_of_unemployment")
```

```{r}
polsmonth_snp_new=left_join(polsmonth_data, snp_data, by =c("year", "month"))
```

```{r}
unemployment_pols_snp_new=left_join(polsmonth_snp_new,unemployment_new_data, by =c("year", "month"))
```
### 
pols-month dataset contains information on , the dataset has 822 observations, 13 variables, the years range from 1947-2015. Some key variables in this dataset include year, month, and presdient.
snp_data dataset contains information on    , the dataset has 787 observations, 3 variables, the years range from 1-12. some key variables include year and month.
unemployment dataset contains information on   , the dataset has 
and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
  
  
  
## problem 3
```{r}
popbabynames_data = read_csv(file = "./data/Popular_Baby_Names.csv")%>% 
  janitor::clean_names()%>%
  mutate(gender=str_to_lower(childs_first_name),
    childs_first_name=str_to_lower(childs_first_name),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = replace (ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
    ethnicity = replace (ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
    ethnicity = replace (ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")
   )%>%
        
   distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank, .keep_all=TRUE)
```


```{r}
olivia_rank=filter(popbabynames_data, gender == "female", childs_first_name == "olivia")
select(ethnicity, rank, year_of_birth)%>%
pivot_wider(names_from = "year_of_birth", values_from = "rank")
```

```{r}
popmale_rank=filter(popbabynames_data, gender == "male")%>%
 select(gender,ethnicity, rank, childs_first_name,year_of_birth)%>%
pivot_wider(names_from = "year_of_birth", values_from = "rank")
```

ethan_pivot=
    filter(popbabynames_data, gender == "MALE", rank =="1")%>%
  select(gender,ethnicity, rank, childs_first_name,year_of_birth)%>%
pivot_wider(names_from = "year_of_birth", values_from = "rank")



```{r}
scatterplot_df=filter(popbabynames_data, gender == "male", year_of_birth=="2016", ethnicity=="white non hispanic")

ggplot(scatterplot_df, aes(x = rank, y = count)) + geom_point()
```

male, white non-hispanic children born in 201

  spread(year, month)

  spread(year, condition, Jan:Feb:Mar:Apr:May:Jun:Jul:Aug:Sep:Oct:Nov:Dec, factor_key=TRUE)

trash_data = janitor::clean_names(litters_data)
    range = "A1:D7")

write_csv(mlb11_data_subset, path = "./data/mlb_subset.csv")
