---
title: "p8105_hw2_yy2912.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
```
## problem 1
## Load in the trash data

```{r}
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel",
                 range= cell_cols("A:N"))%>%
                janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls, digit=0))%>%
  as.integer(sports_balls)

                      
```

```{r}
prec_data_2017=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2017 Precipitation", range = cell_cols("A:B")) %>%
  janitor::clean_names() %>%
  mutate("year"='2017')
```

```{r}
prec_data_2018=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2018 Precipitation", range = cell_cols("A:B")) %>%
  janitor::clean_names() %>%
  mutate(year='2018')
```

```{r}
prec_data_new=  left_join(prec_data_2017, prec_data_2018, by = "Month")
month.name
```
The number of observations in the dataset (trash_data) is 'r nrow (trash_data)'. 

The number of observations in the dataset (prec_data_new) is 'r nrow (prec_data_new)'. 

the key variables in the dataset 

the total precipitation in 2018 

the median number of sports balls in a dumpster in 2017 is 

## problem 2
## Load in the fivethrityeight data

```{r}
polsmonth_data = read_csv(file = "./data/pols-month.csv")%>% 
separate(mon, c("year", "month", "day"))%>% 
mutate(month_numeric=as.numeric(month),month_pols=month.name[month_numeric],president='gop','dem')%>% select(-prez_dem, -prez_gop, -day)

```

```{r}
snp_data = read_csv(file = "./data/snp.csv")%>% 
separate(date, c("year", "month", "day"))%>%
mutate(month=month.name[month_numeric=as.numeric(month)])%>% 
select(-day)%>%
select("year", "month","close")
```

```{r}
unemployment_data = read_csv(file = "./data/unemployment.csv")%>% 
    janitor::clean_names()
```

```{r}
unemployment_new_data = 
pivot_longer(unemployment_data,Jan:Dec,names_to = "month", values_to = "%_of_unemployment")%>%
  janitor::clean_names()
```

```{r}
polsmonth_snp_new=full_join(polsmonth_data, snp_data)
```

```{r}
unemployment_pols_snp_new=full_join(polsmonth_snp_new,unemployment_new_data)
```
### Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
  
  
  
## problem 3
```{r}
popbabynames_data = read_csv(file = "./data/Popular_Baby_Names.csv")%>% 
  janitor::clean_names()%>%
  mutate(childs_first_name=str_to_lower(childs_first_name),)%>%
   distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank, .keep_all=TRUE)
```


```{r}
olivia_rank=filter(popbabynames_data, childs_first_name == "olivia")
```
```{r}
popmale_rank=filter(popbabynames_data, gender == "MALE")
```

```{r}
scatterplot_df=filter(popbabynames_data, gender == "MALE", year_of_birth=="2016", ethnicity=="WHITE NON HISPANIC")

ggplot(scatterplot_df, aes(x = rank, y = count)) + geom_point()
```

male, white non-hispanic children born in 201

  spread(year, month)

  spread(year, condition, Jan:Feb:Mar:Apr:May:Jun:Jul:Aug:Sep:Oct:Nov:Dec, factor_key=TRUE)

trash_data = janitor::clean_names(litters_data)
    range = "A1:D7")

write_csv(mlb11_data_subset, path = "./data/mlb_subset.csv")
