---
title: "p8105_hw2_yy2912.Rmd"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```
## problem 1
## Load in the trash data
```{r}
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel",
                 range= cell_cols("A2:N408"))%>%
                janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls, digit=0), sports_balls_new=as.integer(sports_balls))
```
## read and clean precipitation data from 2017 and 2018
```{r}
prec_data_2017=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate("year"='2017')%>%
  drop_na()

```

```{r}
prec_data_2018=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate("year"='2018')%>%
  drop_na()

```

```{r}
  prec_data_new = 
  bind_rows(prec_data_2017, prec_data_2018) %>%
  mutate(month = month.name [month] )
```

The number of observations in the dataset (trash_data) is `r nrow (trash_data)`

The key variables in the dataset include weight_tons and volume_cubic_yards.

The number of observations in the dataset (prec_data_new) is `r nrow (prec_data_new)` 

The key variables in the dataset include month, total, and year. 

For available data,the total precipitation in 2018 is `r sum(pull(prec_data_2018,total))`

the median number of sports balls in a dumpster in 2017 is `r median(pull(trash_data,sports_balls))`

## problem 2
## Load in the fivethrityeight data
```{r}
polsmonth_data = read_csv(file = "./data/pols-month.csv")%>% 
separate(mon, into=c("year", "month", "day"))%>% 
mutate(month_numeric=as.numeric(month),month_pols=month.name[month_numeric],president=ifelse(prez_gop, 'gop','dem'))%>% 
select(-prez_dem, -prez_gop, -day)

```

```{r}
snp_data = read_csv(file = "./data/snp.csv")%>% 
separate(date, into=c("year", "month", "day"))%>%
mutate(month=month.name[month_numeric=as.numeric(month)])%>% 
select(-day)%>%
select("year", "month","close")
```

```{r}
unemployment_data = read_csv(file = "./data/unemployment.csv")%>% 
    janitor::clean_names() %>% 
  mutate(year=as.character(year))

```  
```{r}
unemployment_new_data = 
pivot_longer(unemployment_data, jan:dec,names_to = "month", values_to = "%_of_unemployment")
```

##join the three cleaned data
```{r}
polsmonth_snp_new=left_join(polsmonth_data, snp_data, by =c("year", "month"))
```

```{r}
unemployment_pols_snp_new=left_join(polsmonth_snp_new,unemployment_new_data, by =c("year", "month"))
```

The pols-month dataset has `r ncol(polsmonth_data)` variables and `r nrow(polsmonth_data)` row. The years range from `r min(pull(polsmonth_data, year))` to `r max(pull(polsmonth_data, year))`. 
The variables in this dataset include `r names(polsmonth_data)`.
Some key variables in this dataset include presdient, gov_gop,sen_gop,rep_gop, and gov_dem,sen_dem,rep_dem. 

The snp_data dataset has `r ncol(snp_data)` variables and `r nrow(snp_data)` rows. The years range from `r min(pull(snp_data, year))` to `r max(pull(snp_data, year))`. 
The variables in this dataset include `r names(snp_data)`.
Some key variables in this dataset include close. 

The unemployment dataset has `r ncol(unemployment_new_data)` variables and `r nrow(unemployment_new_data)` rows. The years range from `r min(pull(unemployment_new_data, year))` to `r max(pull(unemployment_new_data, year))`. 
The variables in this dataset include `r names(unemployment_new_data)`.
Some key variables in this dataset include % of unemployment. 

The merged dataset has `r ncol(unemployment_pols_snp_new)` variables and `r nrow(unemployment_pols_snp_new)` rows.
The range of years in the resulting dataset is from `r min(pull(unemployment_pols_snp_new, year))` to `r max(pull(unemployment_pols_snp_new, year))`.
Some key variables in this dataset include presdient, gov_gop,sen_gop,rep_gop, and gov_dem,sen_dem,rep_dem, close, and % of unemployment. 

  
  
## problem 3
```{r}
popbabynames_data = read_csv(file = "./data/Popular_Baby_Names.csv")%>% 
  janitor::clean_names()%>%
  mutate(gender=str_to_lower(gender),
    childs_first_name=str_to_lower(childs_first_name),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = replace (ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
    ethnicity = replace (ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
    ethnicity = replace (ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")
   )%>%
        
   distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank, .keep_all=TRUE)
```


```{r}
olivia_rank=filter(popbabynames_data, gender == "female", childs_first_name == "olivia")%>%
select(ethnicity, rank, year_of_birth)%>%
pivot_wider(names_from = "year_of_birth", values_from = "rank")
```
 

```{r}
popmale_rank=filter(popbabynames_data, gender == "male", rank == "1")%>%
 select(gender,ethnicity, rank, childs_first_name,year_of_birth)%>%
pivot_wider(names_from = "year_of_birth", values_from = "rank")

```

##create a scatterplot showing the number of children with a name against the rank in popularity of that name for male, white non-hispanic in 2016.
```{r}
scatterplot_df=filter(popbabynames_data, gender == "male", year_of_birth=="2016", ethnicity=="white non hispanic")

ggplot(scatterplot_df, aes(x = rank, y = count)) + geom_point()
```


